% !TEX root = ../notes.tex

\section{January 14, 2019}

Given an interval $(a,b) \subset \R$, we know that the size of this interval is $b-a$. The focus of this course will be the study of the generalization of this idea using the \emph{Lebesgue measure} on $\R$. Equipped with this, we can talk of the \emph{Lebesgue integral} of ``nice'' functions, which is more powerful than the Riemannian equivalent.

\subsection{The Metric Space}

\begin{definition}[Metric Space]
Given a set $X$, a metric function $d$ is a function $d : X \times X \to \R$ obeying the following three properties.
\begin{enumerate}
\item \textbf{Positivity:} $d(x,y) \geq 0$ and $d(x,y) = 0$ if and only if $x = y$;
\item \textbf{Symmetry:} $d(x,y) = d(y,x)$ for all $x,y \in X$;
\item \textbf{Triangle Inequality:} $d(x,y) \leq d(x,z) + d(z,y)$ for all $x,y,z \in X$. 
\end{enumerate}
A metric space is a pair $(X,d)$ where $d$ is a metric function on $X$.
\end{definition}

\begin{example}[Metric Spaces]
\begin{parts}
\part In $\R$, we have the traditional $d(x,y) = \abs{x-y}$.
\part In $\R^2$, we have $d\qty((x_1,x_2),(y_1,y_2)) = \sqrt{(x_1-y_1)^2 + (x_2-y_2)^2}$.
\part In $\R^2$, we also have $d\qty((x_1,x_2),(y_1,y_2)) = \max\{\abs{x_1-y_2}, \abs{x_2-y_2}\}$.
\part The discrete metric on a set $X$ is defined by 
\[ d(x,y) = \begin{cases}
0 & \text{if $x = y$,} \\
1 & \text{otherwise.}
\end{cases} \]
\part Given a metric space $(X,d)$ and $Y \subset X$ then $(Y,d)$ is also a metric space where $d$ is restricted to $Y \times Y$.
\end{parts}
\end{example}

\begin{definition}[Neighborhood]
Fix a metric space $(X,d)$. For some $r \geq 0$, the $r$-neighborhood of $x$ is $B(x,r)$, the set $\{y \in X \mid d(x,y) < r\}$. Notice that this depends on the metric! In $\R$ with the discrete metric, $B(0,1) = \{0\}$ while $B(0,2) = \R$ which is not what we expect from the traditional metric.
\end{definition}

\begin{definition}[Interior Points]
Let $A \subset X$. A point $x \in A$ is an interior point of $A$ if there exists some $r > 0$ such that $B(x,r) \subset A$. That is, we can draw a ball around $x$ which lies entirely in $A$.
\end{definition}

\begin{example}
If $A = [0,1)$, then the interior points of $A$ are $(0,1)$ but $0$ is not an interior point.
\end{example}

\begin{definition}[Open Sets]
A subset $A \subset X$ is open if every point in $A$ is interior. The empty set is vacuously open.
\end{definition}

\begin{proposition}
For any $x \in X$ the $r$-neighborhood of $x$ is an open subset of $X$.
\end{proposition}

\begin{proof}
Let $y \in B(x,r)$. Let $r_0 = r - d(x,y)$. Then $r_0 > 0$ and $B(y,r_0) \subset B(x,r)$ regardless of which $y$ is chosen since for any $z \in B(y,r_0)$ we know that $d(x,z) \leq d(x,y) + d(r,z) < d(x,y) + r_0 < r$. Then every point of $B(x,r)$ is interior and so it is open.
\end{proof}

Using this, we can now call $B(x,r)$ the open ball of radius $r$ centered at $x$.

\begin{example}
In $\R^2$ with the standard metric, an open ball looks like an open disc. With the maximum metric, it looks like an open square. In $\R$, we can look at the set of all rational numbers $\Q$. This set is not open since for all $q \in \Q$ and all $r > 0$ there exists an $x \in B(q,r)$ where $x \notin \Q$.
\end{example}

\begin{proposition}\label{prop:1-2}
The intersection of finitely many open sets is open. The union of any open sets is open.
\end{proposition}

\begin{example}
The intersection of infinitely many open sets is not necessarily open. Consider $\bigcap\, (0,1/n)$ as $n \to \infty$. The intersection is simply $\{0\}$ which is not an open set.
\end{example}

\begin{proof}[Proof of Proposition~\ref{prop:1-2}]
Let $A_1, \cdots, A_k$ be open subsets of $X$. Let $x \in A_1 \cap \cdots \cap A_k$. Since each $A_i$ is open we know that $x$ is an interior point of $A_i$, so there exists some $r_i$ such that $B(x,r_i) \subset A_i$. Let $r$ be the minimum of all such $r_i$. Then $B(x,r) \subset A_i$ for all $i$, so this open ball is contained in the intersection.

Now let $\{A_\alpha \mid \alpha \in I \}$ be a collection of open subsets. Let $x \in \bigcup\,A_\alpha$. Then $x$ is contained in some open $A_\alpha$, and so there exists an $r_\alpha$ such that $B(x,r_\alpha) \subset A_\alpha$, so $B(x,r_\alpha) \subset \bigcup\,A_\alpha$.
\end{proof}

\begin{definition}[Interior of a Set]
For $A \subset X$, the set of all interior points of $A$ is called the interior of $A$. This is usually written as $\operatorname{Int}(A)$ or $A^\circ$.
\end{definition}

\begin{example}
If $A = [a,b]$ then $A^\circ = (a,b)$. If $A = \Q$ then $\Q^\circ = \emptyset$.
\end{example}

\begin{proposition}
For all $A$, the interior of $A$ is open. Furthermore, $A^\circ$ is the largest open subset of $A$ in the sense that it contains all other open subsets of $A$.
\end{proposition}

\begin{proof}
It's just the definitions.
\end{proof}

\begin{proposition}
If $A \subset B$ then $A^\circ \subset B^\circ$.
\end{proposition}

\begin{corollary}
A set $A$ is open if and only if $A = A^\circ$.
\end{corollary}

\begin{definition}[Limit Point]
Let $A \subset X$. A point $x \in X$ is a limit point of $A$ if for any $r > 0$ we know that $B(x,r) \cap A \not= \emptyset$. Notice that every point $a \in A$ is a limit point of $A$.
\end{definition}

\begin{example}
Let $A = [0,1)$. Then $0$ is a limit point of $A$ since every open ball centered centered at $0$ interesects $A$. Furthermore, $1$ is also a limit point for the same reason. If $A = \Q$, then the set of limit points of $\Q$ is all of $\R$.
\end{example}

\begin{definition}[Closed Set]
A set $A \subset X$ is called closed if every limit point of $A$ is contained in $A$.
\end{definition}

\begin{example}
The interval $[0,1]$ is closed but $[0,1)$ is not. To show that something isn't a limit point, use the minimum distance between this point and the interval. This must be positive since otherwise it would be in the interval. Then let your $r$ be smaller than this, and the open ball with this radius centered at this point will not intersect the original interval. Generalize to higher dimensions as needed.
\end{example}

\begin{corollary}
Given any metric space $X$, we know that $\emptyset$ is closed. Furthermore, $\bar{B}(y,r) = B[y,r] = \{y \mid d(x,y) \leq r \}$ is closed for any $r$.
\end{corollary}

\begin{proposition}
Let $A \subset X$. We know that $A$ is open if and only if $A^\complement$ is closed.
\end{proposition}