% !TEX root = ../notes.tex

\section{January 17, 2019}

\epigraph{``The blood gurgled out of his neck as he gasped his last breath.''}{Shelly}

\subsection*{Outline of Lecture}

\begin{enumerate}
\item Application of Morality --- Applied/Practical Ethics
\item Content of Morality --- Normative Ethics
\item Foundation of Morality --- Normative \& Metaethics
\item Nature of Morality --- Normative and Metaethics
\end{enumerate}

\subsection*{Questions of Moral Philosophy}

\begin{example}[Questions of Moral Philosophy]
\begin{parts}
\part What are the various moral rules? What are the specific details? For example, if ``Don't kill'' is the rule, is it every permissible to kill people? When and why? If a child picks up a loaded gun and the only way to stop them from killing people is to shoot the child, is that moral? \emph{(This is called an innocent threat)} Are there proportionality considerations?

\part Do we have obligations to keep our promises? What if we have conflicting promises? How should we weight our promises? Are there ever a valid set of rules we can follow? Can we break our promises if a more pressing situation arises where we could do more good (saving someone in cardiac arrest but skipping a meeting to do so)?

\part Can war ever be moral? What about collateral damage and civilian deaths?

\part Is abortion moral or immoral? Does killing a fetus break the ``don't kill'' command? When do we have rights to life?
\end{parts}
\end{example}

The subfields of morality (although there really isn't a clear line between these areas):

\begin{enumerate}
	\item \textbf{Content of Morality} --- What are the rules of a just society? What factors make an action right or wrong? 
	\item \textbf{Application of Morality} --- How can we apply the rules of morality to real-life situations? What happens when situations are ambiguous? What conflicts arise? Abortion and Capital Punishment? This class doesn't really touch on this.
	\item \textbf{Foundation of Morality} --- Can our rules be derived from one another? Is a prohibition on killing entirely separate from ``Do not harm people'' or is it just a specific case one the more general rule? How to we systematize ethical rules? Is there a single basic rule from which all others derive (like the Golden Rule or Categorical Imperative or Utilitarianism)? What makes the true rules true? Is true even the right word? Why are moral systems just? By virtue of what is morality founded? Divine commandment, contractarianism, etc.
	\item \textbf{Nature of Morality} --- What's ethics all about? What's the point of ethics? What does it even mean to be moral? Why are some things more valid than others? Why do values even exist? How do we include morality in a scientific worldview? Are there ``morality facts''? How could we ever hope to know whether something is right or wrong? What do ``right'' and ``wrong'' even mean? We probably won't touch on this here either (Take Moral Skepticism with Shelly in the fall).
\end{enumerate}